\mytitle

### 2020-04-22 paper.tex
Edited to end of Sect 2.1

We start this section by presenting the physical model of a wind farm
used throughout the paper.  Our wind farm model is based on the common
notation and description presented in~\cite{Zhang2014}.  Subsequently,
we provide a literature review of existing mathematical optimization
methods for solving WFLO, focusing on their similarities and
differences with our approach.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square-shaped wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in squared meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  In this work, we
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we focus on the
novel quadratic approach for WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional wind farm settings.

@@

\paragraph{Proximity Constraints}
Turbines must be placed at least \emph{five rotor diameters}
apart. Depending on the cell size, we shall formulate the
corresponding constraint as part of the optimization model. For
example, assume that a turbine is placed in the center of a
cell. Then, given a cell size of $c = 100$ meters and a rotor diameter
of $40$ meters, we must ensure that two turbines are not placed in
two adjacent cells. In Figure~\ref{fig:field_model}, the red area
around the turbine represents the forbidden cells.

\paragraph{Wakes}
Turbines influence each other through interference effects that are
referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
Wakes cause reduction of the effective power produced by a turbine due
to upstream turbines that change the airflow dynamics. In
Figure~\ref{fig:field_model}, the wind is assumed to be
blowing from left to right. The wake effect due to placing the turbine
expands as shown in the diagram.
%: it will influence only turbines that
%are placed in the orange cells.  In other words,
If a turbine is placed in the orange area, its effective power is
reduced due to the upstream turbine.
% that we placed in the $8\times8$ farm.
Moreover, a given turbine may be placed downstream from multiple turbines and hence
multiple wakes must be taken into account for each location.  The
reduction of power due to the wake effect is determined by three
parameters: (1) the distance between the upstream turbine and the cell
for which we want to compute the effect (denoted by $x$ in
Figure~\ref{fig:field_model}), (2) the radius of the cone's opening
(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
superposition of several wind regimes (a wind regime is a combination
of free speed and directions).  The wake effect can be pre-computed
for each pair of cells since we know: the distance ($x$), the turbine
specifications and terrain conditions (that jointly dictate $r$), and
probabilistic behavior of wind regimes~\cite{Zhang2014}.

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-eastern wind blowing at a free velocity of $12\mbox{ km/h}$. The
probability of this regime is $0.008$.

Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
Note that turbines can adjust their direction according to the current wind 
regime.

Given the interference of the wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. The set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid w.r.t.\ to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}):
\begin{equation}
  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
\end{equation}

  Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes~\cite{jensen1983note}. 
  However, the expression is not amenable to optimization since it is a squared root function of a cubic expression. 
  Hence, a simplified representation for the total expected energy that considers wakes is required.

@@

Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint. \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic. Sounds like it is not natural and we choose it due to its optimization properties. Arik: I changed the arguments - please see the abstract and the beginning of the introduction.]}

>>>>

We start this section by presenting the physical model of a wind farm
%used throughout the paper.
%Our wind farm model is
based on the common
notation and description presented in Zhang et al.\ \cite{Zhang2014}.  Subsequently,
we provide a literature review of existing mathematical optimization
methods for solving WFLO, focusing on their similarities and
differences with our approach.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square-shaped wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  We
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we focus on the
novel quadratic approach for WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional problem characteristics.

@@

\paragraph{Proximity Constraints}
Turbines must be placed at least \emph{five rotor diameters}
apart. Depending on the cell size, we shall formulate the
corresponding constraint as part of the optimization model. For
example, if a turbine is placed in the center of a
cell, a cell size of $c = 100$ meters and a rotor diameter
of $40$ meters, requires that two turbines are not placed in adjacent cells. In Figure~\ref{fig:field_model}, the red area
around the turbine represents the forbidden cells.

\paragraph{Wakes}
Turbines influence each other through \emph{wake} interference effects~\cite{jensen1983note,shakoor2016wake}.
%that are
%referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
Wakes cause reduction of the effective power produced by a turbine due
to upstream turbines that change the airflow dynamics. In
Figure~\ref{fig:field_model}, the wind is assumed to be
blowing from left to right. The wake effect due to placing the turbine
expands as shown in the diagram.
%: it will influence only turbines that
%are placed in the orange cells.  In other words,
If a turbine is placed in the orange area, its effective power is
reduced due to the upstream turbine.
% that we placed in the $8\times8$ farm.
Moreover, a given turbine may be placed downstream from multiple turbines and hence
multiple wakes must be taken into account for each location.  The
reduction of power due to the wake effect is determined by three
parameters: (1) the distance between the upstream turbine and the cell
for which we want to compute the effect (denoted by $x$ in
Figure~\ref{fig:field_model}), (2) the radius of the cone's opening at distance $x$ 
(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
combination of several wind regimes where a wind regime consists of 
ther wind's free speed and direction.  The wake effect can be pre-computed
for each pair of cells since we know: the distance ($x$), the turbine
specifications and terrain conditions (that jointly dictate $r$), and
probabilistic behavior of wind regimes~\cite{Zhang2014}.

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes.
%where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-easterly wind blowing at a free velocity of $12\mbox{ km/h}$. The
probability of this regime is $0.008$.

Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.\footnote{Turbines can adjust their orientation according to the current wind 
regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
Rather, $u_{id}$, can be computed using the following equation~\cite{Zhang2014}:
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. The set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid w.r.t.\ to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines:
%~(\cite{Zhang2014}):
\begin{equation}
  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
\end{equation}

  Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes~\cite{jensen1983note}. 
  However, the expression is challenging to exact mathematical optimization techniques due to its complexity (i.e., the cube of a square root).
  %since it is a squared root function of a cubic expression. 
  %Hence, a simplified representation for the total expected energy that considers wakes is required.

@@

Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
naturally results in a quadratic objective function in our WFLO formulations. Such functions have been the subject of recent research advances \todo{[REFs]}, motivating the contributions here.
%, which is a desired property from an optimization standpoint. \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic. Sounds like it is not natural and we choose it due to its optimization properties. Arik: I changed the arguments - please see the abstract and the beginning of the introduction.]}
